---
title: "Intro to ML - Project (Churn)"
author: "Abhinav, Archit, Ayush, Nir, Matthew"
date: "7/18/2021"
output: html_document
number_sections : true

---
Setting up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#########################
##Cleaning Data and EDA##
#########################
#Install libraries if not installed, else load them-----------------------------
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}
# usage
packages <- c("ggplot2", "ISLR", "DataExplorer", "RColorBrewer", "dplyr", "data.table","rpart","randomForest","xgboost","DescTools","Hmisc","ggcorrplot","MASS","tidyverse","caret","precrec")
ipak(packages)

options(scipen=999)

#Set seed and working directory-------------------------------------------------
set.seed(100)
#setwd("C:/Users/archi/OneDrive/Documents/GitHub/STA380-69963/ML - Project 1")#--WD for Archit
setwd("~/Documents/GitHub/Bank-Customer-Churn-Prediction")#--WD for Abhinav
```


Exploring data
```{r}
#Read data file
raw_data = fread('bank_user_data.csv')


n = dim(raw_data)[1]
df = data.frame(raw_data)

#Understanding the structure of data
str(df)
```

```{r}
#Checking if there are any null value in the dataframe
sapply(df,function(df) sum(is.na(df)))
```

```{r}
#Checking Unique value counts in each columns
sapply(df, n_distinct)

#Exploring unique values for few variables
unique(df$NumOfProducts)
unique(df$HasCrCard)
unique(df$IsActiveMember)
unique(df$Exited)
```


```{r}

## To make data frames easily readable we have 
## removed unnecessary data fields
df = subset(raw_data, select = -c(RowNumber,CustomerId,Surname))
df = data.frame(df)

#Coerce response variable to factor
df$Exited = factor(as.character(df$Exited),levels = c("0","1"))

#Plotting Histograms to understand the distributions
par(mfrow = c(3, 2))

hist(df$CreditScore,main=c("Credit Score"),xlab = c(""),col = "lightblue")
hist(df$Age,main=c("Age"),xlab = c(""),col = "lightblue")
hist(df$Tenure, breaks = 10,main=c("Tenure"),xlab = c(""),col = "lightblue")
hist(df$Balance, breaks = 12,main=c("Balance"),xlab = c(""),col = "lightblue")
hist(df$EstimatedSalary,main=c("Estimated Salary"),xlab = c(""),col = "lightblue")
```
We observe more or less gaussian distribution for credit score and age (with slight right skew) whereas Tenure and Estimated salary more or less are uniform in distribution. A big peak is seen in balance variable distribution at zero implying quite a lot of customers with zero balance accounts.

```{r}
#Plotting Bar Charts to understand the Categorical Variables
theme_custom <- function () { 
    theme_bw(base_size=12, base_family="Avenir") %+replace% 
        theme(
            panel.background  = element_blank(),
            plot.background = element_rect(fill="gray96", colour=NA), 
            legend.background = element_rect(fill="transparent", colour=NA),
            legend.key = element_rect(fill="transparent", colour=NA)
        )
}

ggplot(df, aes(x = factor(Gender))) + geom_bar(fill="skyblue2",alpha=0.65) +xlab("Gender")+
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, colour = "black")+theme_custom()

ggplot(df, aes(x = factor(Geography))) + geom_bar(fill="skyblue2",alpha=0.65) +xlab("Geography")+
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, colour = "black")+theme_custom()

ggplot(df, aes(x = factor(NumOfProducts))) + geom_bar(fill="skyblue2",alpha=0.65)+xlab("Number of Products")+
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, colour = "black")+theme_custom()

ggplot(df, aes(x = factor(HasCrCard))) + geom_bar(fill="skyblue2",alpha=0.65) +xlab("Has Credit Card")+
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, colour = "black")+theme_custom()

ggplot(df, aes(x = factor(IsActiveMember)))+ geom_bar(fill="skyblue2",alpha=0.65)+xlab("Active membership")+
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, colour = "black")+theme_custom()

ggplot(df, aes(x = Exited)) + geom_bar(fill="skyblue2",alpha=0.65) +
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, colour = "black")+
  labs( x = "Exit Status")+theme_custom()

```
We observe majority customers are using credit card and roughly half of them have active membership. Rate of customer exit is 20%.

```{r}
#Customer Churn by Region
#table1 <- table(df$Exited, df$Geography, dnn=c("Exit Count", "Geography")) 
#barplot(table1, ylab="Frequency", xlab="Geography", main="Comparing Exit Status across countries\n", 
#        col=c("turquoise4", "turquoise2" ), beside=TRUE, width=.2)
#legend("right", title="Exited", legend= sort(unique(df$Exited)),
#       fill =c("turquoise4", "turquoise2" ), box.lty=0)

```


```{r}
## Geography vs Gender
cols <- c("Gender","Geography","NumOfProducts","HasCrCard","IsActiveMember","Exited" )
df[cols] <- lapply(df[cols], function(x) as.factor(as.character(x)))

g <- ggplot(df, aes(x = Geography)) +geom_bar(aes(fill = Gender),position="dodge")+geom_text(aes(label = ..count..),position = "identity", stat = "count", vjust =5 , colour = "black")+scale_fill_manual(values=c('#999999','#E69F00'))+theme_custom()
plotly::ggplotly(g+ggtitle("Gender Distribution by Geography"))



```
We see similar gender distributions across all the countries.

```{r}
## Customer Churn by Geography
plotly::ggplotly(ggplot(df, aes(x = Geography, fill = Exited)) +
  geom_bar(position="dodge") +
  geom_text(aes(label = ..count..),
            stat = "Count",position = position_dodge(0.8),
            vjust = 1.5, hjust = 0.5,
            colour = "black") +
  scale_fill_manual(values=c('#999999','#E69F00'))+
  labs(title = "Churn by Geography")+theme_custom())
```

France despite having almost half of the customers from our sample has fewer exits. Germany on the contrary sees more attrition.

```{r}
##Trying to plot in percent---WIP
## Customer Churn by Gender
plotly::ggplotly(ggplot(df, aes(x = Gender,y = (..count..)/sum(..count..), fill = Exited)) +
  geom_bar(position="dodge") +
  geom_text(aes(label =scales::percent( (..count..)/sum(..count..)) ,y = (..count..)/sum(..count..)),
            stat = "Count",position = position_dodge(0.8),
            vjust = 1.5, hjust = 0.9,
            colour = "black") +
  scale_fill_manual(values=c('#999999','#E69F00'))+
  labs(title = "Churn by Gender")+theme_custom())

```
Roughly one in every three women versus one in every five men are exiting the concerned banking services.

```{r}
#Density Plots
plotly::ggplotly(ggplot(df, aes(x=Age)) + 
  geom_density(fill="grey",alpha=0.65)+theme_custom())

plotly::ggplotly(ggplot(df, aes(x=Age,fill=Exited),size=1.3) + 
  geom_density(alpha=0.65)+
  labs(title = "Density Plot: Age")+theme_custom())

#ggplot(df, aes(x=NumOfProducts)) +geom_density()+scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"))+labs(title = "Density Plot: #Products")

#ggplot(df, aes(x=NumOfProducts, color=Exited)) + geom_density()+scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"))+labs(title = "Density Plot: #Products")

```
Age may as well be a good predictor given the observable difference in mean Age for customers retained vs lost.

```{r}
#Correlation Analysis

#str(df)
cols_num <- c("CreditScore","Age","Tenure","NumOfProducts","Balance","EstimatedSalary" )
df[cols_num] <- lapply(df[cols_num], function(x) as.numeric(x))
corr <- round(cor(df[cols_num]), 2)
#ggcorrplot(
#  corr,
#  hc.order = TRUE,
#  type = "lower",
#  outline.color = "white",
#  ggtheme = ggplot2::theme_gray,
#  colors = c("#6D9EC1", "white", "#E46726")
#)
plotly::ggplotly(ggcorrplot(corr,outline.color = "black",colors = c("salmon","white","skyblue2"),lab = T)+ggtitle("Correlation plot\n"))
```

Almost all continuous variables are uncorrelated thereby satisfying absence of multicollinearity assumption of logistic regression.

```{r}
#Check impactables with simple logistic regression
model.glm <- glm(Exited~.,df,family = "binomial")
summary(model.glm)


```
In addition to demographics such as age, gender and location,  actionable variables such as Credit score, Customer balance and active membership are strong predictors. Given above p-values might change as per number of variables, we try running stepwise regression to better understand best combination of features that we can use.

```{r}
step.model <- model.glm %>% stepAIC(trace = FALSE)
cols <- names(coef(step.model))
summary(step.model)

```


```{r}
#Setting up train validation test 
# Split the data into training and test set
set.seed(123)

training.samples <- df$Exited %>%  createDataPartition(p = 0.9,list = F)
train.data  <- df[training.samples, ]
test.data <- df[-training.samples, ]

# Define training control
train.control <- trainControl(## 10-fold repeated CV
                           method = "repeatedcv",
                           number = 10,repeats = 10)
# Train the model
train.data <- train.data[, colnames(train.data) %in% c(cols,"Exited")]
test.data <- test.data[, colnames(test.data) %in% c(cols,"Exited")]

cv_logit <- train(data = train.data, Exited~.,
                 method = "glm", 
                 family = "binomial",
                 trControl = train.control)
summary(cv_logit)

PredTrain = predict(cv_logit, newdata=train.data[,!colnames(train.data) == "Exited"], type="prob")
table(train.data$Exited, PredTrain[,2] > 0.5)

```


```{r fig.width=6.5, fig.height=5.5}
precrec_obj <- evalmod(scores = PredTrain[,2], labels = train.data$Exited)
autoplot(precrec_obj)
#precrec_obj <- evalmod(scores = PredTrain[,2], labels = train.data$Exited,mode="basic")
#autoplot(precrec_obj)

```



```{r}




```
#Matthew Logistic Regression
library(tidyverse)
library(mosaic)

#Read CSV
bank <- read.csv("Churn_Modelling.csv", stringsAsFactors = FALSE)
names(bank)


#Credit Score (.001)
Credit <- glm(Exited ~ CreditScore, data=bank, family="binomial")
summary(Credit)


library(rms)
Credit2 <- lrm(Exited ~ CreditScore, data=bank)
Credit2

#Balance (.022)
Bal <- glm(Exited ~ Balance, data=bank, family="binomial")
summary(Bal)

Bal2 <- lrm(Exited ~ Balance, data=bank)
Bal2

#Age (.115)
Ag <- glm(Exited ~ Age, data=bank, family="binomial")
summary(Ag)

Ag2 <- lrm(Exited ~ Age, data=bank)
Ag2

#Tenure (0)
Ten <- glm(Exited ~ Tenure, data=bank, family="binomial")
summary(Ten)

Ten2 <- lrm(Exited ~ Tenure, data=bank)
Ten2

#EstimatedSalary (0)
Salary <- glm(Exited ~ EstimatedSalary, data=bank, family="binomial")
summary(Salary)

Salary2 <- lrm(Exited ~ EstimatedSalary, data=bank)
Salary2

#NumOfProducts (.004)

tally(~NumOfProducts, data=bank)

bank <- bank %>%
  mutate(NumOfProducts_C = factor(NumOfProducts, labels=c("1", "2", "3", "4")))

Products_C <- glm(Exited ~ NumOfProducts_C, data=bank, family="binomial")
summary(Products_C)

Products_C2 <- lrm(Exited ~ NumOfProducts_C, data=bank)
Products_C2


Products <- glm(Exited ~ NumOfProducts, data=bank, family="binomial")
summary(Products)

Products2 <- lrm(Exited ~ NumOfProducts, data=bank)
Products2

exp(Products_C2$coef)
exp(confint.default(Products_C2))

#HasCrCard (0)
Card <- glm(Exited ~ HasCrCard, data=bank, family="binomial")
summary(Card)

Card2 <- lrm(Exited ~ HasCrCard, data=bank)
Card2

#IsActiveMember (.038)
Active <- glm(Exited ~ IsActiveMember, data=bank, family="binomial")
summary(Active)

Active2 <- lrm(Exited ~ IsActiveMember, data=bank)
Active2

#Geography (.043)
Geo <- glm(Exited ~ Geography, data=bank, family="binomial")
summary(Geo)

Geo2 <- lrm(Exited ~ Geography, data=bank)
Geo2

#Gender (.018)
Gen <- glm(Exited ~ Gender, data=bank, family="binomial")
summary(Gen)

Gen2 <- lrm(Exited ~ Gender, data=bank)
Gen2


#Model (.19)
Full <- glm(Exited ~ Age + IsActiveMember + Geography + Gender, data=bank, family="binomial")
summary(Full)

Full2 <- lrm(Exited ~ Age + IsActiveMember + Geography + Gender, data=bank)
Full2

exp(Full2$coef)
exp(confint.default(Full2))



